SunPy - Python for Solar Physics

### 2014-01-27 4-Util.tex

\begin{listing}[H]
\begin{minted}{pycon}
>>> from sunpy import wcs
Find the helioprojective coordinate of a given longitude and latitude on the Sun
>>> wcs.convert_hg_hpc(10,53)
(100.49244115330731, 767.97438321917502)
Convert that position back to heliographic coordinates
>>> wcs.convert_hpc_hg(100.49, 767.97)
(9.9996521808465175, 52.999563684874893)
Try to convert a position which is not on the Sun to heliographic coordinates
>>> wcs.convert_hpc_hg(-1500,0)
sunpy/wcs/wcs.py:180: RuntimeWarning: invalid value encountered in sqrt
  distance = q - np.sqrt(distance)
(nan, nan)
Convert sky coordinate to a position on the Sun in cartesian coordinates (HCC)
>>> wcs.convert_hpc_hcc(-300,400, z = True)
(-216716967.63331246, 288956420.9477042, 594364636.2208252)
\end{minted}
\caption{Using the sunpy.wcs sub-package.}
\label{code:wcs_code}
\end{listing}

@@

>>> from sunpy.sun import constants as sun
>>> print(sun.mass)
  Name   = Solar mass
  Value  = 1.9891e+30
  Error  = 5e+25
  Units  = kg
  Reference = Allen's Astrophysical Quantities 4th Ed.
Verify the average density of the Sun and convert to cgs
>>> (sun.mass/sun.volume).cgs
<Quantity 1.40851154227 g / (cm3)>
Check to see if the age of the Sun is available and grab the unit and value separately
>>> sun.find('age')
['age', 'average angular size', 'average density', 'average intensity']
>>> sun.value('age'), sun.unit('age')

\end{minted}
\caption{Using the sunpy.sun.constants sub-package.}
\label{code:constants_code}
\end{listing}

\subsubsection{Instruments}
In addition to providing support for instrument-specific solar data via the main data classes \texttt{Map}, \texttt{LightCurve} and \texttt{Spectrum}, some instrument-specific functions may be found within the \texttt{instr} package. These functions are generally those that are unique to one particular solar instrument, rather than of general use, for example a function to construct a GOES flare event list or a function to query the LYRA timeline annotation file. Currently, support is included in this way for the GOES, LYRA, RHESSI and IRIS instruments, while future developments will include support for additional missions. It is anticipated that, ultimately, solar missions requiring a large suite of software tools will be supported via a separately maintained package affiliated with SunPy.

>>>>

\begin{listing}[H]
\begin{minted}{pycon}
>>> from sunpy import wcs
>>> wcs.convert_hg_hpc(10,53)
(100.49244115330731, 767.97438321917502)
# Convert that position back to heliographic coordinates
>>> wcs.convert_hpc_hg(100.49, 767.97)
(9.9996521808465175, 52.999563684874893)
# Try to convert a position which is not on the Sun to HG
>>> wcs.convert_hpc_hg(-1500,0)
sunpy/wcs/wcs.py:180: RuntimeWarning: invalid value encountered in sqrt
  distance = q - np.sqrt(distance)
(nan, nan)
Convert sky coordinate to a position in HCC
>>> wcs.convert_hpc_hcc(-300,400, z = True)
(-216716967.63331246, 288956420.9477042, 594364636.2208252)
\end{minted}
\caption{Using the sunpy.wcs sub-package.}
\label{code:wcs_code}
\end{listing}

@@

>>> from sunpy.sun import constants as sun
>>> print(sun.mass)
  Name   = Solar mass
  Value  = 1.9891e+30
  Error  = 5e+25
  Units  = kg
  Reference = Allen's Astrophysical Quantities 4th Ed.
# Verify the average density of the Sun and convert to cgs
>>> (sun.mass/sun.volume).cgs
<Quantity 1.40851154227 g / (cm3)>
# Search for the age of the Sun
>>> sun.find('age')
['age', 'average angular size', 'average density', 'average intensity']
>>> sun.value('age'), sun.unit('age')
(4600000000.0, Unit("yr"))
\end{minted}
\caption{Using the sunpy.sun.constants sub-package.}
\label{code:constants_code}
\end{listing}

\subsubsection{Instruments}
In addition to providing support for instrument-specific solar data via the main data 
classes \texttt{Map}, \texttt{LightCurve} and \texttt{Spectrum}, 
some instrument-specific functions may be found within the \texttt{instr} package. 
These functions are generally those that are unique to one particular solar instrument, 
rather than of general use, for example a function to construct a GOES flare event list 
or a function to query the LYRA timeline annotation file. Currently, some support is included
 in this way for the GOES, LYRA, RHESSI and IRIS instruments, while future developments 
 will include support for additional missions. It is anticipated that, ultimately, solar
  missions requiring a large suite of software tools will be supported via a separately 
  maintained package affiliated with SunPy.

### 2014-01-27 5-Dev.tex

\section{Development and Community}
SunPy is a community-developed library, designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence 
\footnote{\url{http://opensource.org/licenses/BSD-2-Clause}}, the whole 
development 
process is also online and open for anyone to contribute to.
SunPy's development makes use of the online 
service GitHub \footnote{\url{http://github.com}} and \href{http://git-scm.com/}{git} 
as its distributed version control software.

The continued success of an open source projects depends on many factors of which 
three of the most important are (1) utility and quality of the code, (2) documentation, and (3) an
active community \cite{bangerth2013}. Several tools (some specific to Python) are used by
Sunpy to make achieving these goals more accessible. To maintain high quality code, a 
transparent and collaborative development workflow made possible by Github is used (described
further below). Sunpy includes a number of unit tests which test the functionality 
of the code base. Additionally, Sunpy makes use of `continuous integration' provided by
Travis\footnote{\url{http://travis-ci.org}}, a process by which the addition of any new code 
automatically triggers a comprehensive review of all unit tests. If any single test
fails then the community is alerted before the code is accepted. The unit test coverage is
also monitored by the \href{http://coveralls.io}{Coveralls} service.

Quality documentation is
one of the most important factors determining the success of any software project. 
Thankfully powerful tools already exist in Python to support this thanks to native
Python's focus on its own documentation. Sunpy makes use of the Sphinx\footnote{\url{http://sphinx-doc.org}}
documentation generator. Sphinx uses reStructuredText as its markup language, 
an easy-to-read, what-you-see-is-what-you-get plaintext markup syntax. It supports
many output formats most notably HTML as well as PDF, and ePub and provides a rich
hierarchically structured view of in-code documentation strings. The Sunpy documentation 
s built automatically and hosted on Readthedocs \footnote{\url{http://readthedocs.org}} 
at \url{http://docs.sunpy.org}.

>>>>

\section{Development and Community}
SunPy is a community-developed library, designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence, the whole 
development 
process is also online and open for anyone to contribute to.
SunPy's development makes use of the online 
service GitHub (\url{http://github.com}) and git\footnote{\url{http://git-scm.com/}}
as its distributed version control software.

The continued success of an open source projects depends on many factors of which 
three of the most important are (1) utility and quality of the code, (2) documentation, and (3) an
active community \cite{bangerth2013}. Several tools (some specific to Python) are used by
Sunpy to make achieving these goals more accessible. To maintain high quality code, a 
transparent and collaborative development workflow made possible by Github is used (described
further below). Sunpy includes a number of unit tests which test the functionality 
of the code base. Additionally, Sunpy makes use of `continuous integration' provided by
Travis(\url{http://travis-ci.org}), a process by which the addition of any new code 
automatically triggers a comprehensive review of all unit tests. If any single test
fails then the community is alerted before the code is accepted. The unit test coverage is
also monitored by the \href{http://coveralls.io}{Coveralls} service.

Quality documentation is
one of the most important factors determining the success of any software project. 
Thankfully powerful tools already exist in Python to support this thanks to native
Python's focus on its own documentation. Sunpy makes use of the Sphinx (\url{http://sphinx-doc.org})
documentation generator. Sphinx uses reStructuredText as its markup language, 
an easy-to-read, what-you-see-is-what-you-get plaintext markup syntax. It supports
many output formats most notably HTML as well as PDF, and ePub and provides a rich
hierarchically structured view of in-code documentation strings. The Sunpy documentation 
s built automatically and hosted on Readthedocs (\url{http://readthedocs.org})
at \url{http://docs.sunpy.org}.

### 2014-01-27 solarsoft.tex

It is overwhelmingly an environment which relies upon IDL (Interactive Data 
Language), a commercial and closed-source interactive data analysis environment 
sold by Exelis, a global aerospace, defence, information and services company. 
The SSW `gen' package provides the base analysis capabilities such as time 
series analysis, time conversions, spectral fitting, image display, and file 
I/O. It is composed of 89\% IDL code while the remainder is mostly Perl (5.4\%) 
and csh scripts (3.8\%) \footnote{{\bf This data was calculated using...}}. 
According to data generated using David A. Wheeler's 'SLOCCount' 
\footnote{\url{http://www.dwheeler.com/sloccount/}}, SSW/gen is composed of 
289,724 lines of code with a total approximate value of \$10.4M (based on the 
Basic COCOMO model \citep{_cocomo_2014}). \footnote{{\bf This data was 
calculated using...} If we can add them to the repository will improve the 
quality of the paper and we can always repeat the calculation in the future}. 
Individual missions and instruments provide additional optional packages to 
support and process their data. For example, the SDO package \citep{sdo} is 
100\% IDL code and is 11k lines of code (\$360k) and the SOHO package 
\citep{soho} is 100k lines of code (\$3M) and is 99.9\% IDL code. SSW is hosted 
and distributed by the Lockheed Martin Solar and Astrophysics Laboratory.
\footnote{Do we know whether they have some licence owning it? do we know the 
SolarSoft licence at all?}
%The RHESSI package is 99.97\% IDL code and is 145k total lines of code (\$5M)
While SSW is open-source and freely available, the IDL core is not. The cost 
of IDL is prohibitive for many undergraduates as well as institutions 
with limited means. As well as this the development of SolarSoft is closed, 
with no open issue tracker or community contribution mechanism.

One of SunPy's key aims is to provide a free and modern alternative to the 
SolarSoft (gen) library. This is made possible by the recent rise of the 
scientific Python community. Python, as described in Section \ref{sec:Intro}, 
is a general purpose high-level programming language which is free to use and 
cross-platform and powerful programming language.
It therefore has a very large user community which extends far outside of the 
solar physics or astronomy communities. It is very extensible with C, C++, 
FORTRAN or even IDL. Many books and on-line documentation is available. For 
example at the time writing a search 
%
\footnote{We need to be careful with this! IDL means more things that our 
beloved ``Interactive Data Language''. If we search for that and programming 
then we get just 7 books! - from which 5 are just the same. I don't know how we 
could make a more accurate search.
If we script it it would be amazing!  Also, I would just provide the order
of magnitude, and not a exact number, i.e. 190 and 1800 results} 
%
on Amazon for books on IDL programming generated $189$ results while a search 
for Python yielded $1\,788$ results. Python has a strong emphasis on 
readability a concept which is well-aligned with the scientific endeavour.
Additionally, Python is increasingly being taught in universities, which means 
that research projects can begin quickly without the need for students to learn 
a new (and relatively niche) language. Finally, knowledge of Python is a skill 
which is useful and well compensated outside of the scientific community 
(average salary of \$110k in the Washington DC area according to indeed.com at 
the time of writing). Python continues to see increased use in the astronomy 
community \citep{greenfield2011} which has similar goals and requirements as 
the solar physics community. Finally, Python integrates well with many 
technologies such as web servers \citep{dolgert2008}, SQL databases, 
interactive and shareable notebook-based computing \citep{perez2007}. For all 
of these reasons, Python is an ideal platform for a modern solar physics data 
analysis environment.

>>>>

It is overwhelmingly an environment which relies upon IDL (Interactive Data 
Language), a commercial and closed-source interactive data analysis environment 
sold by Exelis, a global aerospace, defence, information and services company. 
The SSW `gen' package provides the base analysis capabilities such as time 
series analysis, time conversions, spectral fitting, image display, and file 
I/O. It is composed of 89\% IDL code while the remainder is mostly Perl (5.4\%) 
and csh scripts (3.8\%). 
According to data generated using David A. Wheeler's 'SLOCCount'\footnote{\url{http://www.dwheeler.com/sloccount/}}, SSW/gen is composed of 
289,724 lines of code with a total approximate value of \$10.4M (based on the 
Basic COCOMO model \citep{_cocomo_2014}). 
%stuart - This data was calculated using...} If we can add them to the repository 
%will improve the quality of the paper and we can always repeat the calculation in the future.
%schriste - i don't know what you mean by this 
Individual missions and instruments provide additional optional packages to 
support and process their data. For example, the SDO package \citep{sdo} is 
100\% IDL code and is 11k lines of code (\$360k) and the SOHO package 
\citep{soho} is 100k lines of code (\$3M) and is 99.9\% IDL code. SSW is hosted 
and distributed by the Lockheed Martin Solar and Astrophysics Laboratory.
%The RHESSI package is 99.97\% IDL code and is 145k total lines of code (\$5M)
While SSW is open-source and freely available, the IDL core is not. The cost 
of IDL is prohibitive for many undergraduates as well as institutions 
with limited means. As well as this the development of SolarSoft is closed, 
with no open issue tracker or community contribution mechanism.

One of SunPy's key aims is to provide a free and modern alternative to the 
SolarSoft (gen) library. This is made possible by the recent rise of the 
scientific Python community. Python, as described in Section \ref{sec:Intro}, 
is a general purpose high-level programming language which is free to use and 
cross-platform and powerful programming language.
It therefore has a very large user community which extends far outside of the 
solar physics or astronomy communities. It is very extensible with C, C++, 
FORTRAN or even IDL. Many books and on-line documentation is available. For 
example at the time writing a search 
%
%stuart - \footnote{We need to be careful with this! IDL means more things that our 
%beloved Interactive Data Language. If we search for that and programming 
%then we get just 7 books! - from which 5 are just the same. I don't know how we 
%could make a more accurate search.
%If we script it it would be amazing!  Also, I would just provide the order
%of magnitude, and not a exact number, i.e. 190 and 1800 results}
% 
% schriste - I think we defined IDL properly. I was being generous with the book search for IDL
on Amazon for books on IDL programming generated $189$ results while a search 
for Python yielded $1\,788$ results. Python has a strong emphasis on 
readability a concept which is well-aligned with the scientific endeavour.
Additionally, Python is increasingly being taught in universities, which means 
that research projects can begin quickly without the need for students to learn 
a new (and relatively niche) language. Finally, knowledge of Python is a skill 
which is useful and well compensated outside of the scientific community 
(average salary of \$110k in the Washington DC area according to indeed.com at 
the time of writing). Python continues to see increased use in the astronomy 
community \citep{greenfield2011} which has similar goals and requirements as 
the solar physics community. Finally, Python integrates well with many 
technologies such as web servers \citep{dolgert2008}, SQL databases, 
interactive and shareable notebook-based computing \citep{perez2007}. For all 
of these reasons, Python is an ideal platform for a modern solar physics data 
analysis environment.
